A meta-analysis of powdery mildew yield loss mitigation strategies in Australian mungbean

### 2020-07-22 04_MetaAnalysis.Rmd
Updated Emersons meta-analysis file (08) with changes made in the original (04)file

```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management - 1,
   method = "ML",
   random = ~ spray_management | trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```
```{r}
(PM_mv$b - PM_mv$ci.ub) / PM_mv$se
```

In this result we can see that the `Early` treatment is not significantly different from zero, which in this case is the mean of the no spray `control`.
However the other treatments are significantly different from zero.
The $Q_M$ [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, shows the moderators significantly influence the model ($Q_M = 26.22, df = 5, p < 0.0001$) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010].
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 568.31, df = 111, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

Lets look at the trial variability in this model with a forest plot
```{r forestP, fig.height= 40, fig.width=17}
source(here("R/change_labels.R"))
source(here("R/forest_rows.R"))

### Sort and arrange data-frame occordingly and produce a index vector to inform the order of each "Trial" on the plot
index1 <- dat1 %>%
   mutate(ID = seq_along(trial_ref)) %>%
   mutate(spray_management = 
             factor(spray_management, 
                    levels = c("Early","Late","Recommended",
                    "Recommended_plus","Late_plus"))) %>%
   arrange(desc(spray_management), location, desc(grain_MD)) %>%
   mutate(ID2 = order(ID))%>%
   pull(ID2)

### Use forest rows function to return a list of where to place on the plot: sub-titles, each trial row, and summary prediction for each treatment.
### this list can be used below to reference where to plot each variable
rows1 <- forest_rows(head_row = 156, 
                     ord_var = dat1$spray_management, 
                     index = index1, 
                     gap = 5, 
                     row_offset = -2)

### A data-frame of meta-model coefficients used to produce the summary polygons for each treatment
polys <- data.frame(ref = change_labels(names(coef(PM_mv))),
                    x = coef(PM_mv),
                    ci.lb = PM_mv$ci.lb,
                    ci.ub = PM_mv$ci.ub)[c(1,4,2,3,5),]
polys$rows <- rows1$pred_row - 1.5

forest1  <- metafor::forest(PM_mv, cex = 1.15 , width = 5,
                #order = "fit",
                rows = rows1$trows-1.5,
                ilab = cbind(dat1$year, 
                             dat1$row_spacing, dat1$host_genotype, dat1$dose),
                ilab.xpos = c(-2.35,-2.1,-1.8,-1.4),
                slab = NA, 
                xlim = c(-2.9, 2.2),
                ylim = c(18,157),
                xlab = "Grain yield mean difference",
                addfit = FALSE,
                efac = c(0.2,0.15,0.2),
                at = seq(-1,1.5, 0.5))

### code which puts lines and annotations on the plot, its a bit long...
source(here("R/add_annotations.R"))

@@

The first table in this output shows the tau^2 (variance) of each random effects and the number of occurrences for each treatment in the analysis.
The second table is in two parts (left and right). 
The left part, rho, is the correlation of variation between the specified treatments.
All comparisons were acceptable except for a comparison between `Early` and `Late_plus`, `0.000` rho.
`Early` and `Late_plus` treatments never occurred within the same trial, which is indicated by the right side of the table (hence the warning).

>>>>

```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_MD,
   vi,
   mods = ~ spray_management - 1,
   method = "ML",
   random = ~ spray_management | trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```

@@

The first table in this output shows the tau^2 (variance) of each random effects and the number of occurrences for each treatment in the analysis.
This effectively shows the heterogeneity between experiments for these partivular treatments.

The second table is in two parts (left and right). 
The left part, rho, is the correlation of variation between the specified treatments.
All comparisons were acceptable except for a comparison between `Early` and `Late_plus`, `0.000` rho.
`Early` and `Late_plus` treatments never occurred within the same trial, which is indicated by the right side of the table (hence the warning).

### 2020-07-22 08_MetaAnalysis_Emerson.Rmd
Updated Emersons meta-analysis file (08) with changes made in the original (04)file

if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq())

@@

```{r metafor_organisation}
dat1 <-
   slimmer_PM_dat %>%
   mutate(vi = vi/n) %>% 
   #filter(fungicide_ai != "control") %>% #remove controls from the data
   mutate(spray_management =
             factor(
                spray_management,
                c(
                   "control",
                   "Early",
                   "Recommended",
                   "Recommended_plus",
                   "Late",
                   "Late_plus"
                )
             ),
          D_pres = 
             factor(D_pres,
                    c("lowD",
                      "highD")),
          grain_MD = grain_yield.t.ha - yi_C,
          id = row_number()) %>%
   select(
      trial,
      trial_ref,
      location,
      year,
      row_spacing,
      host_genotype ,
      spray_management,
      fungicide_ai,
      D_pres,
      grain_yield.t.ha,
      grain_SMD,
      grain_MD,
      dose,
      vi,
      yield_error,
      vi2,
      n,
      id)
```

@@

```{r}
hist_usq(dat1$grain_yield.t.ha)
hist_usq(log(dat1$grain_yield.t.ha))
hist_usq(sqrt(dat1$grain_yield.t.ha))
hist_usq(dat1$grain_MD)
```

Standardised mean differences have a bit of a long tail however this is not bad and transformations with log or square root will not work on negative values.
So it looks like no transformation is necessary before we start the meta-analysis.

Let's undertake the meta-analysis using the package `metafor`.
We are using the `spray_management` variable as a moderator and an interactive term to the `trial` random variable.
Because we want to know the difference between all treatments and we have no reference treatment, we will remove the intercept.

@@

```{r PM_Mv_Contrast}
source("R/simple_summary.R")
summary(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))), test=adjusted("none"))
simple_summary(summary(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))), test=adjusted("none")))
plot(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))))
```

```{r DisPressComparison}
PM_mv2 <- rma.mv(
   yi = grain_yield.t.ha,
   vi,
   mods = ~ spray_management*D_pres,
   method = "ML",
   random = list(~ spray_management | trial, ~1 | id), 
   struct = "UN",
   control=list(optimizer="optim"),
   data = dat1
)
summary(PM_mv2)
summary(glht(PM_mv2, linfct=cbind(contrMat(rep(7,12), type="Tukey"))), test=adjusted("none"))
plot(glht(PM_mv2, linfct=cbind(contrMat(rep(7,12), type="Tukey"))))

anova(PM_mv, PM_mv2)

@@

Results show with the exclusion of the no spray `control` and `Late` treatments, `Early` applications are significantly worse than all other treatments leading to lower yield estimates.
`Late_plus` is estimated to save the most yield however it is only significantly higher than `Early` and `Late` treatments.
Let's put these values in a plot to visualise the differences between treatments.

First we will format the results into a data frame.

@@

@@

```
To to a better comparison I need to know how the mitigation factor is calculated with the estimated yields

>>>>

if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq())
source(here("R/reportP.R"))

@@

```{r metafor_organisation}
dat1 <-
   slimmer_PM_dat %>%
   mutate(vi = vi/n) %>% 
   mutate(spray_management =
             factor(
                spray_management,
                c(
                   "control",
                   "Early",
                   "Recommended",
                   "Recommended_plus",
                   "Late",
                   "Late_plus"
                )
             ),
          D_pres = 
             factor(D_pres,
                    c("lowD",
                      "highD")),
          grain_MD = grain_yield.t.ha - yi_C,
          id = row_number()) %>%
   dplyr::select(
      trial,
      trial_ref,
      location,
      year,
      row_spacing,
      host_genotype,
      spray_management,
      fungicide_ai,
      D_pres,
      grain_yield.t.ha,
      grain_MD,
      dose,
      vi,
      yield_error,
      vi2,
      n,
      id)
```

@@

```{r}
hist_usq(dat1$grain_yield.t.ha)
hist_usq(log(dat1$grain_yield.t.ha))
hist_usq(sqrt(dat1$grain_yield.t.ha))
hist_usq(dat1$grain_MD)
```
The distribution of data does not seem to be improved by transformation of the data.
We will retain the continuous variable grain_yield.t.ha as the response.
Mean differences won't be used as we are following the methodology of a ARM-based model, where mean differences are calculated following the meta-analysis.

To do a ARM based model we will use the package `metafor`.
We are using the `spray_management` variable as a moderator and an interactive term to the `trial` random variable.

@@

The first table in this output shows the tau^2 (variance) of each random effects and the number of occurrences for each treatment in the analysis.
This effectively shows the heterogeneity between experiments for these partivular treatments.

The tau^2 also gives us the heterogeneity between trials and indicates `Recommended_plus` showed the highest heterogeneity between experiments followed by `Late_plus`, `Late` then `Recommended`, with `Early` and the no-spray `control` showing the least heterogeneity.

The second table is in two parts (left and right). 
The left part, rho, is the correlation of variation between the specified treatments.
All comparisons were acceptable except for a comparison between `Early` and `Late_plus`, `0.000` rho.
`Early` and `Late_plus` treatments never occurred within the same trial, which is indicated by the right side of the table (hence the warning).

In this result we can see that the `Early` treatment is not significantly different to the intercept, which in this case is the mean of the no spray `control`.
However the other treatments are significantly different from the no-spray control (intercept).
The $Q_M$ [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, shows the moderators significantly influence the model ($Q_M =$ `r PM_mv$QM` $, df = $ `r PM_mv$m`, `r reportP(PM_mv$QMp)`) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010].
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = $`r PM_mv$QM` $, df = $`r PM_mv$k - PM_mv$m`, `r reportP(PM_mv$QEp)` ) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

```{r PM_Mv_Contrast}
source("R/simple_summary.R")
summary(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))), test=adjusted("none"))
simple_summary(summary(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))), test=adjusted("none")))
```

These contrasts can be viewed in a plot
```{r plotContrasts}
plot(glht(PM_mv, linfct=cbind(contrMat(rep(1,6), type="Tukey"))))
```

Does disease pressure on the crop affect the optimum spray scenario? 
Above we can see spraying at the first sign of powdery mildew or with in `r min(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"| slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` - `r max(as.numeric(slimmer_PM_dat[slimmer_PM_dat$spray_management == "Late"|                   slimmer_PM_dat$spray_management == "Late_plus","fungicide_timing_1"]))` days after first sign is effective at mitigating yield loss.
Lets test this hypothesis by adding it as a moderator and comparing the new model with the original model above.

```{r DisPressComparison}
PM_mv2 <- rma.mv(
   yi = grain_yield.t.ha,
   vi,
   mods = ~ spray_management*D_pres,
   method = "ML",
   random = list(~ spray_management | trial, ~1 | id), 
   struct = "UN",
   control=list(optimizer="optim"),
   data = dat1
)

anova(PM_mv, PM_mv2)
```
We can see that the lower $QE$ indicates that disease preasure moderator helped explain some of the residual heterogeneity in differences in grain yield between treatments. 
However according to the Wald-test the models were not significantly different and we should use the orriginal "Reduced" model.

The P value was close to the 0.05 cut-off so we will present the model output for the interest of the readers.
```{r DissPressMetMod}
summary(PM_mv2)
summary(glht(PM_mv2, linfct=cbind(contrMat(rep(7,12), type="Tukey"))), test=adjusted("none"))
plot(glht(PM_mv2, linfct=cbind(contrMat(rep(7,12), type="Tukey"))))
```

@@

Lets present the meta-analysis results for the moderator variables in a forest style plot.
First lets create a table of the results.

@@

### Forest plot of trial treatment effect per trial
Lets look at the trial variability in this model with a forest plot
```{r forestP, fig.height= 40, fig.width=17}
source(here("R/change_labels.R"))
source(here("R/forest_rows.R"))

### Sort and arrange data-frame accordingly and produce a index vector to inform the order of each "Trial" on the plot
index1 <- dat1 %>%
   mutate(ID = seq_along(trial_ref)) %>%
   mutate(spray_management = 
             factor(spray_management, 
                    levels = c("Early","Late","Recommended",
                    "Recommended_plus","Late_plus"))) %>%
   arrange(desc(spray_management), location, desc(grain_MD)) %>%
   mutate(ID2 = order(ID))%>%
   pull(ID2)

### Use forest rows function to return a list of where to place on the plot: sub-titles, each trial row, and summary prediction for each treatment.
### this list can be used below to reference where to plot each variable
rows1 <- forest_rows(head_row = 156, 
                     ord_var = dat1$spray_management, 
                     index = index1, 
                     gap = 5, 
                     row_offset = -2)

### A data-frame of meta-model coefficients used to produce the summary polygons for each treatment
polys <- data.frame(ref = change_labels(names(coef(PM_mv))),
                    x = coef(PM_mv),
                    ci.lb = PM_mv$ci.lb,
                    ci.ub = PM_mv$ci.ub)[c(1,4,2,3,5),]
polys$rows <- rows1$pred_row - 1.5

forest1  <- metafor::forest(PM_mv, cex = 1.15 , width = 5,
                #order = "fit",
                rows = rows1$trows-1.5,
                ilab = cbind(dat1$year, 
                             dat1$row_spacing, dat1$host_genotype, dat1$dose),
                ilab.xpos = c(-2.35,-2.1,-1.8,-1.4),
                slab = NA, 
                xlim = c(-2.9, 2.2),
                ylim = c(18,157),
                xlab = "Grain yield mean difference",
                addfit = FALSE,
                efac = c(0.2,0.15,0.2),
                at = seq(-1,1.5, 0.5))

### code which puts lines and annotations on the plot, its a bit long...
source(here("R/add_annotations.R"))

@@

```
To to a better comparison I need to know how the mitigation factor is calculated with the estimated yields.
